KidsAIR: exposure model

### 2020-05-01 collinearity_exposure_analysis.Rmd
exposure analysis updates

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(home, area, sampling_day, starts_with("home_act"))

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  left_join(sums_temp_change, by = c("area", "home")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home")) %>% 
  mutate(area = as.factor(area))
```

@@

```{r, eval=TRUE, include=TRUE}
# Filter data for specific variables; select pm variable
pm_mean_sampling_period <- analysis_data %>% 
  select(pm_mean_sampling_period, area, income_3level, 
         residents_smoke,
         home_floors_2level, 
         stove_age_3level, chimney_clean_3level, home_mold,
         wood_collect_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_smoking, 
         home_act_door, home_act_sweep,
         stove_grade_3level) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(pm_mean_sampling_period)

# Filter data for specific variables; select predictor variables
other_vars <- analysis_data %>% 
  select(pm_mean_sampling_period, area, income_3level, 
         residents_smoke,
         home_floors_2level, 
         stove_age_3level, chimney_clean_3level, home_mold,
         wood_collect_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_smoking, 
         home_act_door, home_act_sweep,
         stove_grade_3level) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(-pm_mean_sampling_period)

# Use `omcdiag` to assess overall multicollinearity
omcdiag(other_vars, pm_mean_sampling_period)

# Use `imcdiag` to assess multicollinearity between individual variables
imcdiag(other_vars, pm_mean_sampling_period)

# Specify a method to assess
imcdiag(other_vars, pm_mean_sampling_period, method = "VIF")
```

Initially assessed the following variables for collinearity, based on simple
associations with mean PM over the sampling period:
         pm_mean_sampling_period, area, income_3level, total_residents_2level,
         race_parent_2level, residents_smoke,
         home_type_2level, home_floors_2level, home_windows_2level,
         stove_age_3level, chimney_clean_3level, home_damp, home_mold,
         home_fireplace,
         wood_collect_2level,
         moisture_closest_2level,
         temp_max_5,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_gas, home_act_smoking, home_act_windows, 
         home_act_door, home_act_sweep,
         stove_grade_3level

@@

Based on results and potential overlap, the following vars were removed:
         race_parent_2level, home_damp, moisture_closest_2level,
         temp_max_5, home_act_gas, total_residents_2level, home_type_2level, 
         home_windows_2level, home_fireplace, home_act_windows

@@

Initial variables to use for model selection (14):
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, stove_age_3level,
                      chimney_clean_3level, home_mold, wood_collect_2level
Home Activity: home_act_door, home_act_smoking, home_act_sweep
Others: sums_events_sampling_period, mean_temp_sampling_period_5, stove_grade_3level

>>>>

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>%
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         amb_pm_24hr:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(treatment == "Placebo") %>% 
  #filter(winter_id == 1) %>% 
  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
  group_by(area, home, home_winter_id) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home,  home_winter_id, .keep_all = TRUE)

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  group_by(home) %>% 
  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
                                   lead(residents_smoke), residents_smoke)) %>% 
  ungroup() %>% 
  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
                                 labels = c("<3", "3+"),
                                 right = FALSE))
```

@@

```{r, eval=TRUE, include=TRUE}
# Filter data for specific variables; select pm variable
pm_mean_sampling_period <- analysis_data %>% 
  select(pm_mean_sampling_period, income_3level, 
         residents_smoke,
         home_floors_2level, 
         home_pets_2level,
         chimney_clean_3level, 
         stove_cert, 
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         amb_pm_sampling_period_5,
         home_act_smoke_2level, 
         home_act_door_2level,
         stove_grade_3level) %>% 
  #mutate(area = factor(area)) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(pm_mean_sampling_period)

# Filter data for specific variables; select predictor variables
other_vars <- analysis_data %>% 
  select(pm_mean_sampling_period, income_3level, 
         residents_smoke,
         home_floors_2level, 
         home_pets_2level,
         chimney_clean_3level, 
         stove_cert, 
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         amb_pm_sampling_period_5,
         home_act_smoke_2level, 
         home_act_door_2level,
         stove_grade_3level) %>% 
  #mutate(area = factor(area)) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(-pm_mean_sampling_period)

# Use `omcdiag` to assess overall multicollinearity
#omcdiag(other_vars, pm_mean_sampling_period)

# Use `imcdiag` to assess multicollinearity between individual variables
#imcdiag(other_vars, pm_mean_sampling_period)

# Specify a method to assess
imcdiag(other_vars, log(pm_mean_sampling_period), method = "VIF")
```

Initially assessed the following variables for collinearity, based on simple
associations with mean PM over the sampling period:
         pm_mean_sampling_period, income_3level, area,
         residents_smoke,
         home_floors_2level, 
         home_pets_2level,
         stove_age_3level, chimney_clean_3level, 
         stove_cert, home_mold,
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         amb_pm_sampling_period_5,
         home_act_smoke_2level, 
         home_act_door_2level,
         stove_grade_3level,
         race_parent_2level, home_damp, 
         temp_max_5, home_act_gas, home_type_2level, 
         home_windows_2level, home_fireplace

@@

This is the first sign that `area` has overlap with the other covariates that
will be difficult to deal with. Removing `area` lowers the VIF for several 
of the variables.

Based on results and potential overlap, the following vars were removed:
         race_parent_2level, home_damp, home_mold, stove_age_3level,
         temp_max_5, home_act_gas, home_type_2level, 
         home_windows_2level, home_fireplace

@@

Initial variables to use for model selection (14):
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
                      chimney_clean_3level, wood_collect_2level, 
Home Activity: home_act_door_2level, home_act_smoke_2level
Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,

### 2020-05-01 exposure_models.Rmd
exposure analysis updates

@@

Initial variables to use for model selection (14) from collinearity RMD:
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, stove_age_3level,
                      chimney_clean_3level, home_mold, wood_collect_2level
Home Activity: home_act_door, home_act_smoking, home_act_sweep
Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
        stove_grade_3level, moisture_closest_2level,

Note: sums_events_sampling_period and mean_temp_sampling_period_5 have a lot of
missing data and reduce the sample size from 93 to 43. Not considering these
variables for model selection for now.

Also try adding other variables in to check their impact on model selection:
   total_residents_2level (seems impactful), home_type_2level (not impactful), 
   home_windows_2level (seems impactful), home_act_windows (not very impactful),
   education_3level (not impactful)

Some vars not important in any models, so removing from consideration:
   home_mold, home_floors_2level, home_act_windows, home_type_2level

Next steps:
Some vars (like residents_smoke), other demographics, and home activity vars
have quite a bit of missing data. This is in addition to the missing ambient
and SUMs data. The first step should be to account for as much of this as possible.
See if there are ways to pull demographic data from Winter 2.
Then, see if data for WMT is more complete in general. The idea is to minimize
missing data for model selection and the final model to have as much power as
possible. 
Consider doing a more "full" model for WMT only (if there is less missing data),
particularly since the ambient data is more reliable for WMT than the other areas.
Then you could justify not doing a mixed model, and model selection would be easier.
Then, using data from all areas, use the full model from WMT to work on a 
mixed model that is likely reduced due to sparse data from the other areas.
Since model selection is difficult with mixed models, use the WMT model to inform
this model, while also doing sensitivity analysis to check other variables.
Aside from the multivariable model, much of the paper could also be looking at
simple associations between PM and some of the novel covars (SUMS, ambient data,
stove grades, wood moisture), as well as methods for collecting and analysing 
those vars.
Also look into predict() function for teach/test and validating the model.
Discussion can focus on what important predictors of indoor PM are in these homes - 
if simple demographic data predicts as well as complex things like SUMs and wood
moisture, or if there is a spectrum of predictors and complexity in their collection.

@@

Looking at model selection and averaging after multiple imputation

```{r}
library(mice)
library(pan)
library(multilevel)

# Select analysis vars
selected_vars <- analysis_data %>% 
  select(area, pm_mean_sampling_period, income_3level, 
         stove_age_3level, home_windows_2level,
         chimney_clean_3level, 
         home_act_door_2level, home_act_windows_2level, home_act_smoke_2level,
         sums_events_sampling_period, mean_temp_sampling_period_5, 
         amb_pm_sampling_period_5, stove_grade_3level, moisture_closest_2level) %>% 
  mutate(area = factor(area, levels = c("AK", "NN", "WMT")),
         log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
  select(-pm_mean_sampling_period) %>% 
  filter(!is.na(income_3level) & !is.na(home_windows_2level)) %>% 
  filter(area == "WMT") %>% 
  select(-area)

# Look at structure of data and check variable types
str(selected_vars)

# Check % missing data
mean(is.na(selected_vars))

# Check ICC 
multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))

# Impute missing data; m = 20 datasets (check % missing data and match)
imputed_data <- mice(selected_vars, m=5, maxit = 5,
                     method = c("polr", "polr", "logreg", "polr", 
                                "logreg", "logreg", "logreg", "pmm", "pmm",
                                "pmm", "polr", "logreg", "pmm"))

>>>>

**Note**

Add some of the variables created below to the full dataset

@@

**Step 1**

The model selection process begins with simple associations between mean
PM2.5 over the 6-day sampling period and covariates in the dataset. See
the `exposure_covariate_associations.rmd` file for this process.

Variables that are associated with PM2.5 are selected and assessed for collinearity
in the `collinearity_exposure_analysis.rmd`. Variables that have high overlap
are selected out of the pool of potential covariates to parse down the options;
however, these variables can be used in sensitivity analyses later.

Initial variables to use for model selection (14):
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
                      chimney_clean_3level, wood_collect_2level, 
Home Activity: home_act_door_2level, home_act_smoke_2level
Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,

**Missing data notes**
Some vars (like residents_smoke), other demographics, and home activity vars
have quite a bit of missing data. This is in addition to the missing ambient
and SUMs data. The first step should be to account for as much of this as possible.
See if there are ways to pull demographic data from Winter 2.
Then, see if data for WMT is more complete in general. The idea is to minimize
missing data for model selection and the final model to have as many observations as
possible. 
Consider doing a more "full" model for WMT only (if there is less missing data),
particularly since the ambient data is more reliable for WMT than the other areas.
(ambient temp/PM data for WMT is based on geolocated lat/long, whereas NN and AK
are based on zipcode and have much less spatial variation).
Then you could justify not doing a mixed model, and model selection would be easier.
Then, using data from all areas, use the full model from WMT to work on a 
mixed model that is likely reduced due to sparse data from the other areas.
Since model selection is difficult with mixed models, use the WMT model to inform
this model, while also doing sensitivity analysis to check other variables.
Aside from the multivariable model, much of the paper could also be looking at
simple associations between PM and some of the novel covars (SUMS, ambient data,
stove grades, wood moisture), as well as methods for collecting and analysing 
those vars.
Also look into predict() function for teach/test and validating the model.
Discussion can focus on what important predictors of indoor PM are in these homes - 
if simple demographic data predicts as well as complex things like SUMs and wood
moisture, or if there is a spectrum of predictors and complexity in their collection.

**Looking at model selection and averaging after multiple imputation**

After assessing many model selection possibilities, using multiple imputation
and model selection from imputed datasets seems to be the best option.
There is not a huge amount of missing data, but it is spread out enough
that it eliminates a larger percentage of the homes from consideration for
model selection. Multiple imputation gives us a full dataset to work with,
and there are documented techniques of using this data for model selection.

First, select analysis variables
```{r}
library(mice)
#library(pan)
#library(multilevel)

# Select analysis vars
selected_vars <- analysis_data %>% 
  filter(area == "WMT") %>% 
  select(pm_mean_sampling_period, income_3level, 
         residents_smoke, home_floors_2level, home_pets_2level,
         chimney_clean_3level, stove_cert, wood_collect_2level,
         moisture_closest_2level, sums_events_sampling_period,
         mean_temp_sampling_period_5, amb_pm_sampling_period_5,
         home_act_smoke_2level, home_act_door_2level, stove_grade_3level) %>% 
  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
  select(-pm_mean_sampling_period) 
```

**Note**

When selecting data above, WMT has 63 observations
When use `na.exclude()` to only use complete data (what is needed for AIC selection),
the observations drops to 30. While there is only 10% missing data, it is spread
out enough that excluding it to have complete observations means we can only use
less than half for model selection. This would be a very biased version of model
selection to exclude that much data, which justifies multiple imputation to 
impute the 10% missing data and use that for model selection.
The missing observations are even more extreme for NN and AK.

Next, assess data structure and patterns of missing data
Document missing data thoroughly for manuscript
```{r}
# Look at structure of data and check variable types
str(selected_vars)

# Check % missing data
mean(is.na(selected_vars)) # 10%

# Check ICC 
multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))
```

Next, use `mice()` to impute the missing data
Assess imputed data for summary stats and convergence
```{r}
# Impute missing data; m = 10 datasets (check % missing data and match)
imputed_data <- mice(selected_vars, m=5, maxit = 5,
                     method = c("polr", "logreg", "logreg", "logreg", "polr",
                                "polr", "logreg", "logreg", "pmm", "pmm",
                                "pmm", "logreg", "logreg", "polr", "pmm"))

Below are other options/ideas for model selection
Not using for now, but keep for resources

@@
